Learning Excursion Sets of Vector-valued Gaussian Random Fields for Autonomous Ocean Sampling

### 2021-03-12 main_web.tex
Updates from Overleaf

\begin{figure}[!b] 
\centering 
\subfigure[The waypoint graph.]{\includegraphics[width =
0.49\textwidth]{Figures/web/wp_graph_paper.jpeg}\label{fig:wp_graph_a}}
\hfill
\subfigure[The waypoint graph in 3D.]{\includegraphics[width =
0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
trajectory choices over the $31\times31$ grid used to discretize the GRF. The AUV is set to start in node $53$.
\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment.}
\label{fig:wp_graph}
\end{figure}

Each strategy is conducted on an equilateral triangular grid as shown in
Fig.~\ref{fig:wp_graph}. The AUV starts at the center coordinate at
the southern end of the domain (node 53). It then moves along edges in
the waypoint graph while collecting data which are assimilated onboard
to update the GRF model. This is used in the evaluation of the next
node to sample.  The procedure is run for $10$ stages. A total of 100
replicate simulations were conducted with all strategies.

@@

\begin{figure}[!h]
  \centering
  \subfigure[IBV.]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/web/avg_EV.jpeg}}
  \hfill \subfigure[RMSE between estimated field and
  truth.]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
  \hfill \subfigure[Explained variance
  $\bR^{2}$.]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
  \hfill \subfigure[Computational time for inference (the lines for \textit{naive}, \textit{static\_north}, \textit{static\_east}, and \textit{static\_zigzag}
  overlap).]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
  \caption{Simulation results from 100 replicate simulations for 10
    sampling choices/stages on the grid. Vertical lines show variation
    in replicate results.}
\label{fig:sim_results}
\end{figure}

@@

Fig.~\ref{fig:avg_time} shows the computational effort: the
\textit{naive} strategy is on par with the static designs, while the
\textit{myopic} strategy is slower because it evaluates expected
values for all candidate directions at the waypoints. But it is still
able to do so in reasonable time, which allows for real-world
applicability. The \textit{look-ahead} strategy is much slower, reaching levels that
are nearly impractical for execution on an AUV. Some pruning of the
graph is performed to improve the performance, such as ruling out
repeated visitations. Further pruning of branches or inclusion of
other heuristics could be included for better performance. Then again,
the inclusion of such heuristics is likely a contributing factor for
the \textit{look-ahead} strategy failing to outperform the
\textit{myopic} strategy.

We studied the sensitivity of the results by modifying the input
parameters to have different correlations between temperature and
salinity, standard deviations, and spatial correlation range.  In all
runs, the \textit{myopic} and \textit{look-ahead} strategies perform
the best in terms of realized IBV, and much better than
\textit{naive}. The \textit{look-ahead} strategy seems to be
substantially better than the \textit{myopic} design only for very
small initial standard deviations or very large spatial correlation
range. 
We also ran simulation studies with only temperature data, and for
realistic correlation levels between temperature and salinity, the IBV
results are not much worse when only temperature data are
available. In addition to the comparison made in
Table~\ref{tab:sim_rhoab}, the current setting includes spatial
correlation and this likely strengthen the effect of temperature information. However, it seems that having temperature data alone
does a substantially worse job in terms of explained variance.

@@

The statistical model parameters were specified based on experience as well as data gathered in a short
preliminary survey where the AUV made an initial transect to determine
the trends in environmental conditions, variance and correlation
structures. This gives a reasonable idea of the temperature and salinity of river and ocean waters. Both temperature and salinity were assumed to increase linearly with the west coordinate. Next, empirical correlation and variograms of the residuals from the regression analysis were analyzed to specify the temperature and salinity correlation, variances and spatial correlation decay parameter of the GRF model.
The resulting
parameters are given in Table \ref{tab:experiment_param}. The
regression parameters shown here are scaled to represent the east and
west boundaries of the domain as seen in the preliminary transect
data, and the thresholds are intermediate values. These parameter
values were then used in field trials where we explored the
algorithm's ability to characterize the river plume front separating
the river and fjord water masses.

@@

\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/web/auv1_es_sal_ep.jpeg}\label{fig:res1}}
\hspace{0.2cm}
\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/web/auv4_es_sal_ep.jpeg}\label{fig:res2}}
\caption{Results from mapping the Nidelva river, Trondheim, Norway
  over two survey missions. \ref{fig:map} shows an overview of the
  survey area overlaid with the AUV path in black and dashed
  line. Note the shaded region indicating a typical frontal
  region. \ref{fig:res_both} shows the collected temperature data as
  colored trails. Note waypoint 5 (WP5) which indicates where the two
  surveys diverge. \ref{fig:res1} and \ref{fig:res2} shows the
  collected salinity data overlaid on the final EP, which indicate the
  AUVs statistical impression of the front. For both missions the
  temperature and salinity data correspond with an indication of the
  EP front. About 2 hours time separated the two runs.}
\label{fig:results}
\end{figure*}

@@

Two survey missions (1 and 2), were run successively, with a short
break in between. The resulting path of the selected waypoints are
shown in the map in Fig.~\ref{fig:map}, both within the expected
frontal region (shaded light-red). The recorded temperatures are shown as
colored trails in Fig.~\ref{fig:res_both}, clearly indicating the
temperature difference between fjord and riverine waters. The salinity
data are then shown separately, overlaid with the estimated EP for
each survey in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2}.

Both surveys successfully estimated and navigated the separation zone,
crossing the frontal boundary multiple times. As conditions changed
slightly between the two surveys, the resulting trajectory (after
waypoint 5) is shown to deviate. Survey 1 continued northwards,
tracking the north-eastern portion of the front, while Survey 2 turned
west, mapping the south-western region.

The final predictions of the front location, represented by
conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
dashed lines, correspond with one another. In both surveys they yield
a picture of the front being to the west in the southern portions of
the region and gradually bending off toward the north east. The amount
of exploration done by Survey 1 which turned north is greater than
Survey 2 which was coming close to the survey area borders in the
south-western corner. The AUV paths in both runs are relatively
winding in the sense that the vehicle reacts to the data. In our
experience from simulations, the paths tend to be smoother when we
use larger measurement error variances for the AUV temperature and salinity observations. 
%But here the estimated nugget from the actual trial data is close to the specified level.

\begin{figure}[!b]
  \centering
  \subfigure[Residual plot.]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
  \hfill
  \subfigure[Quadratic form CDFs.]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram.]{\includegraphics[height = 0.252\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
  \caption{Data analysis from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
    along with specified Gaussian contours. \ref{fig:parest_b} Empirical CDF
    (solid) of the quadratic form of the residuals along with the
    modeled theoretical CDF (dashed) of the $\chi^2$ distribution with two
    degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the survey
    salinity and temperature data.}
\label{fig:parest}
\end{figure}

Fig.~\ref{fig:parest} summarizes with diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity
and temperature are subtracted from the data. This scatter-plot of
joint residuals indicates that the correlation between the two
variables might be slightly larger than the initially specified $0.5$ correlation,
and that initial uncertainty in temperature was set a bit too high. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate they should be
approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical
and theoretical curves are not very different, which indicates that
the specified Gaussian model fits reasonably
well. Fig.~\ref{fig:parest_c} shows the empirical variograms of the
scaled residuals for temperature and salinity. The decay is similar
for the two, and seems to be negligible after about $150$ m as was
specified. The working assumption of a separable covariance function
is hence not unreasonable. There appears to be a slight nugget effect, but not much different from what was set as the measurement noise variance.
Overall, the spatial Gaussian model representation is a simplification
of the changing water conditions, and even within our working assumptions,
the parameter specification could likely have been
improved. However, we do not expect that minor changes in parameters would have caused much
difference in AUV sampling paths.

@@

This work builds on a multidisciplinary effort combining statistical
methods with robotic sampling for oceanographic applications. We show
how observation practices can gain efficiency and accuracy from
statistical techniques for spatial monitoring and demonstrate the the
need for real-time multivariate spatial sampling on autonomous
platforms.

@@

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, which would be relevant on a larger
time scale. We do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be a point of departure
\citep{sigrist2015stochastic,richardson2017sparsity}. The sampling
capabilities are restricted by the sensors onboard the AUV, the
computational load, and navigation resources. The sampling of
temperature and salinity variables can be much higher than the grid
resolution, but, due to computational and navigation resources
available on the AUV, assimilation into a discretized fixed grid is
necessary. For more complex oceanographic phenomena, the methods will
need to be extended to non-Gaussian phenomena, possibly feature-based
mixtures of Gaussian processes which could potentially be run onboard
augmented by dynamical models. Running numerical models onboard a
robotic vehicle is currently infeasible, but high-resolution ocean
models or remote sensing data can be used to fit a more complex
statistical model \citep{davidson19}.

The spatio-statistical design criterion building on random sets is
relevant in our setting with different water properties.  We show
mathematical generality beyond the expected integrated Bernoulli
variance, for instance, that of volume uncertainties which is possibly
more relevant, but one that requires more computational resources.
Such criteria could be particularly useful in other oceanographic
settings related to mapping of algal-blooms, anoxic zones or open
water fronts \citep{costa19}.  Other criteria could also be relevant,
for instance, hybrid or multi-attribute criteria that could balance
goals of exploration and exploitation in this situation. Equally, such
techniques have significant use cases in downstream decision-making,
with policy makers and regulators who need to make difficult decisions
related to aquaculture or other marine resources. Value of information
analysis \citep{Eidsvik:15} could be used to evaluate whether
information is likely to result in improved decision-making, in such a
context.  We also foresee opportunities related to design of
experiments for multivariate processes using our notion of generalized
locations.

In our context the myopic strategy performs well, and due to
computational constraints we did not go in depth on the dynamic
programming solutions. There has been much work on finite horizon
optimization in the robotics literature including probabilistic road
maps and rapidly-exploring random trees \citep{karaman2011sampling},
but their statistical properties are unclear.  In some cases it is
also limiting to use a waypoint graph, and would be beneficial to
allow more continuous updates and navigation at the highest frequency
possible given limitations being onboard an AUV.  It is equally
interesting to explore the additional flexibility that can be gained
by having multiple vehicles co-temporally exploring a spatial or
spatio-temporal domain \citep{ferreira2019advancing}. Such an approach
would enable concurrent sampling in different parts of the space, or
opportunities to move in parallel to best capture the excursion set.
The value of information related to when and what to communicate (to
shore or to other vehicles) is also an interesting thrust for research
and likely to be useful for internet-of-things applications or
computer experiments where some observations or evaluations are rather
inexpensive, while others must only be done when they are really
valuable.

>>>>

\begin{figure}[!b] 
\centering 
\subfigure[The waypoint graph]{\includegraphics[width =
0.49\textwidth]{Figures/web/wp_graph_paper.jpeg}\label{fig:wp_graph_a}}
\hfill
\subfigure[The waypoint graph in 3D]{\includegraphics[width =
0.49\textwidth]{Figures/sim/wp_graph_3d.pdf}\label{fig:wp_graph_b}}
\caption{\ref{fig:wp_graph_a} The equilateral waypoint graph used to discretize the
trajectory choices over the $31\times31$ grid used to discretize the GRF. The AUV is set to start in node $53$.
\ref{fig:wp_graph_b} The waypoint grid shown in a 3D environment}
\label{fig:wp_graph}
\end{figure}

Each strategy is conducted on an equilateral triangular grid as shown in
Fig.~\ref{fig:wp_graph}. The AUV starts at the center coordinate at
the southern end of the domain (node 53). It then moves along edges in
the waypoint graph while collecting data which are assimilated onboard
to update the GRF model. This is used in the evaluation of the next
node to sample.  The procedure is run for $10$ stages. In this experiment, a total of 100
replicate simulations were conducted with all strategies.

@@

\begin{figure}[!h]
  \centering
  \subfigure[IBV]{\label{fig:avg_ev}\includegraphics[height=0.49\textwidth]{Figures/web/avg_EV.jpeg}}
  \hfill \subfigure[RMSE between estimated field and
  truth]{\label{fig:avg_rmse}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_RMSE.pdf}}
  \hfill \subfigure[Explained variance
  $\bR^{2}$]{\label{fig:avg_r2}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_R2.pdf}}
  \hfill \subfigure[Computational time for inference (the lines for \textit{naive}, \textit{static\_north}, \textit{static\_east}, and \textit{static\_zigzag}
  overlap)]{\label{fig:avg_time}\includegraphics[height=0.49\textwidth]{Figures/sim/avg_Time.pdf}}
  \caption{Simulation results from 100 replicate simulations for 10
    sampling choices/stages on the grid. Vertical lines show variation
    in replicate results.}
\label{fig:sim_results}
\end{figure}

@@

Fig.~\ref{fig:avg_time} shows the computational effort: the
\textit{naive} strategy is on par with the static designs, while the
\textit{myopic} strategy is slower because it evaluates expected
values for all candidate directions at the waypoints; however, it is still
able to do so in reasonable time, which allows for real-world
applicability. The \textit{look-ahead} strategy is much slower, reaching levels that
are nearly impractical for execution on an AUV. Some pruning of the
graph is performed to improve the performance, such as ruling out
repeated visitations. Further pruning of branches or inclusion of
other heuristics could be included for better performance. Then again,
the inclusion of such heuristics is likely a contributing factor for
the \textit{look-ahead} strategy failing to outperform the
\textit{myopic} strategy.

We studied the sensitivity of the results by modifying the input
parameters to have different correlations between temperature and
salinity, standard deviations, and spatial correlation range.  In all
runs, the \textit{myopic} and \textit{look-ahead} strategies perform
the best in terms of realized IBV, and much better than
\textit{naive}. The \textit{look-ahead} strategy seems to be
substantially better than the \textit{myopic} design only for very
small initial standard deviations or a very large spatial correlation
range. 
We also ran simulation studies with only temperature data, and for
realistic correlation levels between temperature and salinity; the IBV
results are not much worse when only temperature data are
available. In addition to the comparison made in
Table~\ref{tab:sim_rhoab}, the current setting includes spatial
correlation, which likely strengthened the effect of temperature information. However, it seems that temperature data alone
perform substantially worse in terms of explained variance.

@@

The statistical model parameters were specified based on experience as well as data gathered in a short
preliminary survey where the AUV made an initial transect to determine
the trends in environmental conditions, variance and correlation
structures. This provided a reasonable idea of the temperature and salinity of river and ocean waters. Both temperature and salinity were assumed to increase linearly with the west coordinate. Next, empirical correlation and variograms of the residuals from the regression analysis were analyzed to specify the temperature and salinity correlation, variances, and spatial correlation decay parameter of the GRF model.
The resulting
parameters are given in Table \ref{tab:experiment_param}. The
regression parameters shown here are scaled to represent the east and
west boundaries of the domain as seen in the preliminary transect
data, and the thresholds are intermediate values. These parameter
values were then used in field trials where we explored the
algorithm's ability to characterize the river plume front separating
the river and fjord water masses.

@@

\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/web/auv1_es_sal_ep.jpeg}\label{fig:res1}}
\hspace{0.2cm}
\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/web/auv4_es_sal_ep.jpeg}\label{fig:res2}}
\caption{Results from mapping the Nidelva river (Trondheim, Norway)
  over two survey missions. \ref{fig:map} shows an overview of the
  survey area overlaid with the AUV path in black and dashed
  lines. Note the shaded region indicating a typical frontal
  region. \ref{fig:res_both} shows the collected temperature data as
  colored trails. Note waypoint 5 (WP5), which indicates where the two
  surveys diverge. \ref{fig:res1} and \ref{fig:res2} show the
  collected salinity data overlaid on the final EP, which indicate the
  AUV's statistical impression of the front. For both missions, the
  temperature and salinity data correspond with an indication of the
  EP front. About 2 hours time separated the two runs.}
\label{fig:results}
\end{figure*}

@@

Two survey missions (1 and 2) were run successively, with a short
break in between. The resulting path of the selected waypoints are
shown in the map in Fig.~\ref{fig:map}, both within the expected
frontal region (shaded light-red). The recorded temperatures are shown as
colored trails in Fig.~\ref{fig:res_both}, clearly indicating the
temperature difference between fjord and riverine waters. The salinity
data are then shown separately, overlaid with the estimated EP for
each survey in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2}.

Both surveys successfully estimated and navigated the separation zone,
crossing the frontal boundary multiple times. As conditions changed
slightly between the two surveys, the resulting trajectory (after
waypoint 5) also deviated. Survey 1 continued northwards,
tracking the north-eastern portion of the front, while Survey 2 turned
west, mapping the south-western region.

The final predictions of the front location, represented by
conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
dashed lines, correspond with one another. In both surveys they yield
a picture of the front being to the west in the southern portions of
the region and gradually bending off toward the north east. The amount
of exploration done by Survey 1, which turned north, is greater than
Survey 2, which came close to the survey area borders in the
south-western corner. The AUV paths in both runs are relatively
winding in the sense that the vehicle reacts to the data. In our
experience from simulations, the paths tend to be smoother when we
use larger measurement error variances for the AUV temperature and salinity observations. 
%But here the estimated nugget from the actual trial data is close to the specified level.

\begin{figure}[!b]
  \centering
  \subfigure[Residual plot]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
  \hfill
  \subfigure[Quadratic form CDFs]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram]{\includegraphics[height = 0.252\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
  \caption{Data analysis from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
    along with specified Gaussian contours. \ref{fig:parest_b} Empirical CDF
    (solid line) of the quadratic form of the residuals along with the
    modeled theoretical CDF (dashed line) of the $\chi^2$ distribution with two
    degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the survey
    salinity and temperature data.}
\label{fig:parest}
\end{figure}

Fig.~\ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity
and temperature are subtracted from the data. This scatter plot of
joint residuals indicates that the correlation between the two
variables might be slightly larger than the initially specified $0.5$ correlation,
and that initial uncertainty in temperature was set a bit too high. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate, they should be
approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid line) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed line). The empirical
and theoretical curves are not very different, which indicates that
the specified Gaussian model fits reasonably
well. Fig.~\ref{fig:parest_c} shows the empirical variograms of the
scaled residuals for temperature and salinity. The decay is similar
for the two, and seems to be negligible after about $150$ m, as was
specified. The working assumption of a separable covariance function
is therefore not unreasonable. There appears to be a slight nugget effect, though it is not much different from what was set as the measurement noise variance.
Overall, the spatial Gaussian model representation is a simplification
of the changing water conditions, and even within our working assumptions,
the parameter specification could likely have been
improved. However, we do not expect that minor changes in parameters would have caused much
difference in AUV sampling paths.

@@

This work builds on a multidisciplinary effort combining statistical
methods with robotic sampling for oceanographic applications. We show
how observation practices can gain efficiency and accuracy from
statistical techniques for spatial monitoring and demonstrate the
need for real-time multivariate spatial sampling on autonomous
platforms.

@@

The case study considers the upper water column in the river plume,
represented by a two-dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, they would be relevant on a larger
time scale. We do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be a point of departure
\citep{sigrist2015stochastic,richardson2017sparsity}. The sampling
capabilities are restricted by the sensors onboard the AUV, the
computational load, and navigation resources. The sampling of
temperature and salinity variables can be much higher than the grid
resolution, but, due to computational and navigation resources
available on the AUV, assimilation into a discretized fixed grid is
necessary. For more complex oceanographic phenomena, the methods will
need to be extended to non-Gaussian phenomena, possibly feature-based
mixtures of Gaussian processes which could potentially be run onboard and
augmented by dynamical models. Running numerical models onboard a
robotic vehicle is currently infeasible, but high-resolution ocean
models or remote sensing data can be used to fit a more complex
statistical model \citep{davidson19}.

The spatio-statistical design criterion building on random sets is
relevant in our setting with different water properties.  We show
mathematical generality beyond the expected integrated Bernoulli
variance, for instance, that of volume uncertainties which is possibly
more relevant, but one that requires more computational resources.
Such criteria could be particularly useful in other oceanographic
settings related to mapping of algal-blooms, anoxic zones or open
water fronts \citep{costa19}.  Other criteria could also be relevant,
for instance, hybrid or multi-attribute criteria that could balance
goals of exploration and exploitation in this situation. Similarly, such
techniques have significant use cases in downstream decision-making,
with policy makers and regulators who need to make difficult decisions
related to aquaculture or other marine resources. Value of information
analysis \citep{Eidsvik:15} could be used to evaluate whether
information is likely to result in improved decision-making in such a
context.  We also foresee opportunities related to design of
experiments for multivariate processes using our notion of generalized
locations.

In our context the myopic strategy performs well, and due to
computational constraints, we did not go in-depth on the dynamic
programming solutions. There has been much work on finite horizon
optimization in the robotics literature, including probabilistic road
maps and rapidly-exploring random trees \citep{karaman2011sampling},
but their statistical properties are unclear.  In some cases it is
also limiting to use a waypoint graph, and would be beneficial to
allow more continuous updates and navigation at the highest frequency
possible given limitations being onboard an AUV.  It is equally
interesting to explore the additional flexibility that can be gained
by having multiple vehicles co-temporally exploring a spatial or
spatio-temporal domain \citep{ferreira2019advancing}. Such an approach
would enable concurrent sampling in different parts of the space, or
opportunities to move in parallel to best capture the excursion set.
The value of information related to when and what to communicate (to
shore or to other vehicles) is also an interesting avenue for future research
and likely to be useful for internet-of-things applications or
computer experiments where some observations or evaluations are rather
inexpensive, while others must only be done when they are really
valuable.
